---
title: "GR5206 Homework 1"
author: "Hanao Li"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Part 1: Importing Data into R##
```{r}
# i.)
# Import the dataset into R and called titanic
titanic <-read.table('Titanic.txt', header =TRUE, as.is = TRUE)

# ii.)
# check the number of rows and columns titanic have
nrow(titanic)
ncol(titanic)
```
#### So there are 891 rows and 12 columns.

```{r}
# iii.)
# Create a new variable called Survived.Word from the variable 'Survived'
titanic['Survived.Word'] <- NA
titanic$Survived.Word[titanic$Survived == 1] <- 'survived'
titanic$Survived.Word[titanic$Survived == 0] <- 'died'
```
##Part 2: Exploring the Data in R##
```{r}
# i.)
lists <- c('Survived', 'Age', 'Fare')
small <- titanic[,names(titanic) %in% lists]
apply(small,2,mean)
```
#### From the results, the mean of Survived tells us that the survival rate for the Titanic passengers is about 0.38

```{r}
summary(small$Age)
```
#### The mean of Age has a value of NA and this is because there are 177 missing values for the 'Age' variable

```{r}
# ii.)
# Compute the proportion of female passengers who survived the titanic disaster
round(sum(titanic$Sex == 'female' & titanic$Survived == 1) / sum(titanic$Sex == 'female'), 2)
```
#### The proportion is 0.74

```{r}
# iii.)
# Compute the proportion of female passengers out of the survivors
round(sum(titanic$Sex == 'female' & titanic$Survived == 1) / sum(titanic$Survived == 1), 2)
```
#### The proportion is 0.68

```{r}
# iv.)
# Compute the survival rate for each class using for loop
classes <- sort(unique(titanic$Pclass))
Pclass.Survival <- vector('numeric', 3)
names(Pclass.Survival) <- classes

for (i in 1:3){
  Pclass.Survival[i] <- round(sum(titanic$Pclass == i & titanic$Survived == 1) / sum(titanic$Pclass == i), 2)
}
Pclass.Survival
```
#### The survival rate is 0.63, 0.47 and 0.24 for Class 1, Class 2 and Class 3, respectively.

```{r}
# v.)
# Compute the survival rate for each class using tapply
Pclass.Survival2 <- vector('numeric', 3)
names(Pclass.Survival2) <- classes

Pclass.Survival2 <- round(tapply(titanic$Survived, titanic$Pclass, function(x){sum(x) / length(x)}), 2)
Pclass.Survival2
```
#### The survival rate is 0.63, 0.47 and 0.24 for Class 1, Class 2 and Class 3, respectively.

```{r}
# vi.)
# Summary
```
#### From the results we had, we could tell that there is a relationship between survival rate and class. The bette the Class, the higher the survival rate. Class 1 had the highest survival rate while Class 3 had the lowest survival rate.

